Agenda:

Introduction:
1. Purpose of Interview - Zweck des Interviews:
-untersuchen, wie verschiedene xAI Techniken auf den Nutzer wirken und wie diese verbessert werden können.
-Wichtig ist, dass wir uns auf die Erklärung des AI Modells beziehen, nicht das AI Modell an sich (zeigen wir gleich)
2. Terms of Confidentiality - Vertraulichkeit:
-deine Daten werden vertraulich behandelt und anonymisiert verarbeitet. Ausschnitte des Interviews werden in der Veröffentlichung einfließen
-your data will be treated confidently and is anonymized after the interview process. Excerpts from the interview will be used for a scientific submission on explainable AI.
3. Format of the Interview - Interviewformat:
-wir führen ein Semistrukturiertes Interview und orientieren uns an dem xAI Fragebogen
4. Any more questions?

User Profiling:
-answer survey questions
-Professional position, specialisation, experience with ki-67 workflow
-go into detail with personal background if present

Main Image:
-initial impression (of ki-67) -> how do you rate trustworthiness (evaluate whether you can trust the model)
-Wichtig ist, dass wir uns auf die Erklärung des AI Modells beziehen, nicht das AI Modell an sich (zeigen wir gleich)

xAI Survey:
Open questions relating to each of the main criteria:
\begin{itemize}
    \item How do you interpret the explanation shown here?
    \item What does the explanation tell you about how the model is reaching this outcome?
    \item How does the explanation affect your trust in the model output?
    \item How useful could an explanation like this be for you in an AI-assisted workflow?
    \item What could make this type of explanation better?
\end{itemize}
- how do you interpret this explanation, what does it tell you?
 - what is good (about this particular example, about this class of annotation)
 - what could be better, inc. other implementations of the same class/'type' of explanation/annotation
 
 Good Questions we could use again:
-How does this technique fit into your workflow?
-Which is your preferred method 
-How could this help you identify a bad algorithm?