 15:04 Introduction
 15:04 About Interviewee
 Works in Lung Pathology for +10years. AI familiar, but no practical application. Only one experimental solution that does not work. 
 Other paper with ML for Melanimoa? to find histological chars. AI application to evaluate images on smartphone. Want to get feedback what are chars - intensity of color, morpohology, patterns of nuclei.
 
 15:16 Initial Image
 after quick look, I can say I agree with machine. Are all unmarked cells really lymphocytes? Probably not, have to check that. 
 
 15:22 Saliency Map
 Not easy to understand, different colors what do they mean? (not really relevant since it is UI problem). T: What does it tell you about factors? L: Red high intensity nuclei, nucleoulus is also important, this is probably why it
 misses other cells where nucleolus is not as pronounced.
 For global: now I ask myself why are other brown cells not marked? Since I dont see this, I am not satisfied with explanation. 
 T: Is there a better way? L: Not right now, but first image was easiest to understand
 
 15:29 Concept Attribution
 Okay, but now almost too much info, have to read this. My expectation for AI is that is have to save me time. You cant just go after N size, regularity, color, but there are macrophages which dont differ in this way.
 Better applicable not for counting AI, but for tasks such as highlighting tumor tissue.
 
 15:34 CounterFactuals
 I like them since they are straightforward. Interprete this as what degree of staining intensity is taken into account. Only look at positivity/negativity since I remember tumor cells from previous steps, where I already looked for
 other chars 
 
 15:45 TrustScores
 Good, basic, shows us where problem areas were. Almost too easy. Add 3rd color for negative. T: Is there a danger in a solution that provides stuff just at a glance? L: Wont trust software until I have seen it work
 in real life on my own images. Then, I have no problem with simple explanations
 
 15:49 Prototypes
 Still prefer CF, dont get any useful information since I know he can do easiest job. Can he also do weak positive staining? Threshold is more important. T: Where could it be useful L: Only where there are two possible solutions.
 There is a case where only strong result is positive, everything else should be disregarded. The worst xAI so far? Dont get info about density, intensity,... 
 
 15:53 Closing Discussion
 T: What would it take for you to trust it? L: All the data that was used to come to this result? (he means chars/features) plus validation.  C: How would that be structured?
 L: Tested on 
 
 Notes:
 -has completed survey
 -general low scores in some results since animosity for AI