 13:02 Introduction
 13:05 Director of Vivantus Pathology. Using Digital Pathology (old-school), trying CognitionMaster Suite from Charite and WSI. Want to use AI applications since Pathologist Hours are scarce.
 Thoughts about AI application: want to use AI solution, but cant use cloud-based service because of data protection laws.
 T: Primary Barriers? H: Barrier at the moment, have to convince pathologist today by showing good examples which cant happen if we cant apply them. Benefit is in entirely digital workflow, avoid stacks of slides.
 Also, responsibility problem. I want to train my own application (feed annotated pictures). Own work: create synoptic report (paired statement, length and measure, statement and yes).
 
 13:17 First Image
 I want to have usual Agent-E stain, then other stain, then the filtered image. 
 
 13:23 Saliency Map
 What are limits for giving relevance to a nucleus? I want know how do you decide whether a nucleus is deemed pos/neg. I go from low to high magnification, I make up my mind about percentage of cancer cells.
 Other examples might be interesting.
 
 13:30 Concept Attribution
 Ex makes sense, I find it understandable. I feel like I can grasp this, would apply this in the same way - feel more confident in using this since I can see it is using same criteria.
 Improve it by showing limits - when is nucleus pos/neg? Nice to manipulate factors, to see how proportion of cells changes. Program should point out round vs distorted nuclei - visualization of slide examples.
 How are unassociated areas stained?
 
 13:36 Prototypes
 Here you are just talking about staining, not about size and shape. You would need an array of prototypes to visualize this. 
 
 13:43 Counterfactual
 Grid is good, move threshold to improve result. Take control data from the slide? For therapeutic antibodies, we have multiple tissue arrays to evaluate different features?
 
 13:49 Trust Scores
 It is a good idea for the end, after manipulating threshold you can now make a diagnosis, since you can see confidence. Is the app doing the right thing on the control (different cases)?
 I understand that main factor is staining intensity, size is not a stringent requirement - since I can see small cells that are considered highly positive. 
 
 13:55 Discussion
 How would you explain? Highlight, have some text, representative magnification. 
 T: Synoptic Reports H:? Am convinced that SR is a good tool to not forget anything of importance, trust this as colleague.
 
 
 Notes:
 -prostate cancer (cylinder?) as example for AI task
 -project to generate synoptic reports from slides
 -how do you decide whether a nucleus is deemed pos/neg
 -key piece is who was involved in training data - this does not guarantee reliable result!
 -(multidimensional) wheel/field of different features and their interaction
 -how would you communicate/explain relevant factors to others? (per mail)
 -CF divisive: individual threshold, thats why we should set it first and then see how result looks like - cant work as definitive result, I have to tell the truth 
 
 general points:
 -what about thresholds?
 -should this be changeable?
 -strong pos/neg examples misleading
 -saliency maps doesnt tell me anything about factors, only region
 - AI doesnt tell you truth, it just gives you an idea