Liron PANTANOWITZ, Prof. Dr. Department of Pathology, University of
Pittsburgh, PA, USA
lpantanowitz@gmail.com

https://www.pathology.med.umich.edu/faculty/lironp

-expert in clinical pathology with fellowships in hematopathology and cytopathology, leader of pathology field
-leader in informatics with special interest in digital pathology
-commitment for education

1. He is knowledgeable
2. He is commited for education and explanations
3. He is interested since we provide interesting stuff
4. Good feedback

Reasoning:
Prof. Dr. Pantanowitz is perfectly situated at the intersection of pathology and informatics, with extensive knowledge on both fields. His expert knowledge and experience in the field of clinical pathology enables deep insights into the applied field of our study, while his informatics background allows a qualified discussion of the technical background. His commitment for education is also valuable on the account of making the core concepts of medical concepts tangible, and should interest him in our proposed approaches for how future xAI solutions could be designed. We expect Dr. Pantanowitz to find our study highly interesting, since it provides to our knowledge the first qualitative and quantitative survey of xAI solutions in his field, and look forward to his expert feedback from the perspective of pathology.

Claes LUNDSTROEM, Prof. Dr. Department of Science and Technology,
Linköping University, SE
claes.lundstrom@liu.se

https://liu.se/en/employee/clalu03

-focus on medical imaging, at crossroads of ML, visualizaiton, HCI in clinical settings
-uncertainty visualization
-digital pathology research

Mr. Lundström complements the expertise from Dr. Pantanowitz, being an expert in the field of computer science. His background at the intersection of Machine Learning, visualization and Human Computer Interaction perfectly fits the setting of our paper, it being deeply connected to all of these three topics. Mr. Lundströms expert knowledge is further complemented by his engagement in the field of Medical Imaging and especially Pathology, which provides him with the necessary background for a deeper understanding of the needs of our target group. We also expect Mr. Lundström to be interested in the insights of our paper and the conclusions drawn, and expect valuable feedback from his perspective of machine learning and human computer interaction.